---
title: "Quality control on cells and genes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## QC and data process

the number of transplantation cells we had used for this step was more than 2879. This is the reason the cell number remain on this step is not equally to 2301, the number of cells retained for further analysis. 

### Quality control on cells and genes for homeostasis and transplantation cells respectively



```{r}

### load raw cell and meta data
rm(list=ls())
pwd <- getwd()
pwd <- substr(pwd,1,nchar(pwd)-8)
load(paste0(pwd,"/input/01.RawCell_UMI_TPM_Meta.Rdata")) 
```
there are three data tables: raw_cells, raw.counts and raw.tpm:\
raw_cells contains the information of 2879 cells.
```{r}
dim(raw_cells)
```
```{r echo=FALSE}
knitr::kable(head(raw_cells),caption="raw_cells")

```
The raw.counts and raw.tpm are the genes(23154) expression of the cells(2879).
```{r}
dim(raw.counts)
dim(raw.tpm)
```
### Create the output dir
```{r}
dir.create(paste0(pwd,"/output"))
```
### homeostasis data
```{r}
metadata.Ho <- raw_cells[raw_cells$Ho_Tx == "Ho",]
cells.idx.Ho <- match(rownames(metadata.Ho), colnames(raw.counts))
all.counts.Ho <- as.data.frame(raw.counts[, cells.idx.Ho])
dim(all.counts.Ho) # 23154  1569
 
```

### Transpantation data
```{r}

metadata.Tx <- raw_cells[raw_cells$Ho_Tx == "Tx",]
cells.idx.Tx <- match(rownames(metadata.Tx), colnames(raw.counts))
all.counts.Tx <- as.data.frame(raw.counts[, cells.idx.Tx])
dim(all.counts.Tx) # 23154  1310
```
### Quality contol 
#### function 1
SQCstat function can get the total counts of the cell and the total expressed genes of the cell.
```{r}
SCQCstat <- function(data=data, is.expr=1){
  total_counts <- colSums(data, na.rm = T)
  total_genes <- apply(data, 2, function(x){
    sum(x>=is.expr)
  })
  return(list("total_counts"=total_counts, "total_genes"=unlist(total_genes)))#unlist ???æœªçœ‹æ‡‚ï¼ŒæŒ‰é“ç†æ­¤å¤„è¿”å›çš„å‘é‡å€¼ï¼Œå½“ç„¶unlistå¯ä»¥åœ¨è¿”å›åˆ—è¡¨æ—¶æå–å‡ºé‡Œé¢çš„å‘é‡
}

```
#### function 2 
QC function is the main function for HO and Tx quality contol
```{r}
QC <- function(x){   #åŸæ¥è¿™é‡Œæœ‰ä¸ªlabelæ²¡ç”¨çš„å˜é‡ï¼Œåˆ äº†
     if(x=="Ho"){
	 counts.input <- all.counts.Ho  
     meta.input <- metadata.Ho
	 } else {
	 counts.input <- all.counts.Tx
     meta.input <- metadata.Tx
	 }
     # first qc:quality control for cells and genes
     counts.qc <- SCQCstat(counts.input)
     #get the useless data per phenotype
     size.drop <- scater::isOutlier(counts.qc$total_counts, nmads=3, type="both", log=F, batch = meta.input$phenotype)
     gene.drop <- scater::isOutlier(counts.qc$total_genes, nmads=3, type="both", log=F, batch = meta.input$phenotype)
     counts.filter <- counts.input[, !(size.drop | gene.drop)]
     #second qc:filtering with gene number and umi counts
     counts.filter.qc <- SCQCstat(counts.filter)
     # set the threshold value of the express gene number and the counts number
     gene.count <- 1000
     umi.count <- 20000
     counts.filter <- counts.filter[, counts.filter.qc$total_counts >= umi.count & counts.filter.qc$total_genes >= gene.count]
     dim(counts.filter) # Ho:23154 1270; Tx:23154  1059
      # last qc:filtering low abundance genes
     counts.filter.qc <- SCQCstat(counts.filter)
     ave.cut <- 0.2
     ave.counts <- rowMeans(counts.filter)
     counts.filter <- counts.filter[ log10(ave.counts) >= log10(ave.cut),]#æ­¤å¤„è¿™ä¸ª0.2å¹¶æ²¡æœ‰è¿‡æ»¤ä»»ä½•åŸºå›?
  #save data
	 if(x=="Ho"){
     filter.counts.Ho <- counts.filter
     filter.tpm.Ho <- raw.tpm[row.names(counts.filter), colnames(counts.filter)]
     meta.Ho <- meta.input[colnames(counts.filter),]
	 counts.Ho <- counts.input[,colnames(counts.filter)]
	 tpm.Ho <- raw.tpm[,colnames(counts.filter)]
	 save(filter.tpm.Ho, filter.tpm.Ho, counts.Ho, tpm.Ho, meta.Ho, file = paste0(pwd,"/input/01.Homeostasis.Cells.UMI_TPM_metadata.RData"))
	 return(meta.Ho)
	 } else {
     filter.counts.Tx <- counts.filter
     filter.tpm.Tx <- raw.tpm[row.names(counts.filter), colnames(counts.filter)]
     meta.Tx <- meta.input[colnames(counts.filter),]	 
	 counts.Ho <- counts.input[,colnames(counts.filter)]
	 tpm.Ho <- raw.tpm[,colnames(counts.filter)]
	 save(filter.counts.Tx, filter.tpm.Tx, counts.Ho, tpm.Ho, meta.Tx,file = paste0(pwd,"/input/01.Transplantation.Cells.UMI_TPM_metadata.RData"))  
	 return(meta.Tx)
	 }
}

```
#### QC
```{r}
meta.Ho <- QC("Ho")
meta.Tx <- QC("Tx")
## homeostasis cells remain 
dim(meta.Ho) # 1270   10
## transplantation cells remain 
dim(meta.Tx) # 1059   10 
```
