---
title: "Classify transplantation cells"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Classification of transplantation cells to homeostasis cell types
  Note that we used more than 1,031 transplantation cells as training data set to perform this step, the result have been slightly influenced by the number of training cells, so we can see a few cells changed their cell type. However, it didn't affect the conclusions. 

### Set up the environment and load the data
```{r}
rm(list=ls())
pwd <- getwd()
pwd <- substr(pwd,1,nchar(pwd)-8)
### load expression and meta table of homeostasis and transplantation 
load(paste0(pwd,"/input/01.QC.Comb.cell.count.tpm.meta.Rdata"))
### load PCA top genes and clusters of homeostasis cells
load(paste0(pwd,"/input/02.Homeostasis.Cells_cluster_pc.topGenes.RData"))
### load dimensionality reduction results produced by the forward step 
load(paste0(pwd,"/input/04.Comb_tsne.homeostasis_transplantation.RData"))
```

### Devide the data into two groups:homeostasis(train data),transplantation(test data)
```{r}
homeostasis.train <- comb.data.tpm[pc.genes$Gene, row.names(comb.data.meta)[comb.data.meta$Ho_Tx == "Ho"]]
transplantation.test <- comb.data.tpm[pc.genes$Gene, row.names(comb.data.meta)[comb.data.meta$Ho_Tx == "Tx"]]
dim(homeostasis.train) # 1044 1270
dim(transplantation.test) # 1044 1031
```

### Get the factor of true classfications if training data
```{r}
Cell_type.v2 <- paste0("t", pc.cells$Celltype, pc.cells$Cluster)
Cell_type.v2.order <- unlist(sapply(unique(pc.cells$Celltype), function(x){
  paste0("t", x, names(table(pc.cells$Cluster[pc.cells$Celltype == x ])))
}), use.names = F)
Cell_type.v2 <- Cell_type.v2[match(colnames(homeostasis.train), pc.cells$Cell)]
```

### Start knn training
```{r}
# KNN classification
library(class)
# nearest number of cells
Knum <- 10
# by sub cell type
transplantation.classification <- knn(train = t(homeostasis.train), test = t(transplantation.test), 
                                cl = Cell_type.v2, prob = T, k = Knum)
```

### Prepare data frame for transplantation cells
```{r}
# get the tsne map of 2301 cells
tsne.df <- data.frame( x= comb.tsne$Y[,1], y=comb.tsne$Y[,2])
# combine tsne map with cell information
tsne.df <- cbind(tsne.df, comb.data.meta)
tsne.df$predict_cell_type <- c(Cell_type.v2, as.character(transplantation.classification))
tsne.df$predict_cell_type.family <- tsne.df$predict_cell_type
tsne.df$predict_cell_type.family <- factor(tsne.df$predict_cell_type.family, levels=Cell_type.v2.order, ordered=T)
time.order <- c("Ho", "1Day", "3Day", "5Day", "7Day")
tsne.df$Time <- factor(tsne.df$Time, levels = time.order, ordered = T)

```

### Save data 
```{r}
save(tsne.df,file = paste0(pwd,"/input/05.Cluster.tsne.homeostasis_transplantation.RData"))
```





